---
title: "Target Rate Impacts Measure of Discriminability (d')"
runtime: shiny
output: html_document
---

[Trisha N. Patel](https://tnpatel.github.io/) 

Signal Detection Theory (SDT) is used in situations where one is trying to detect a signal among noise. The SDT framework provides a measure of discriminability (d') that can summarize the ability to accurately distinguish information-baring patterns from random patterns. 

The following simulation demonstrates how target rate (the number of targets presented in a set of trials) can influence the measurement of d'.


### Parameter Inputs

```{r echo = FALSE}
fluidRow(
  
  column(4,
         wellPanel(
           sliderInput("dpop", label = "Population d'",
                       min = 0, max = 3, value = 0.7, step = .1),
           sliderInput("crit", label = "Criterion",
                       min = -3, max = 3, value = 0, step = 0.1),
           sliderInput("ntrials", label = "Number of Trials",
                       min = 1, max = 500, value = 100, step = 1),
           radioButtons("nsim", label= "Number of Simulations in a Sample", choices = c("100", "1000"), inline= TRUE)
           
         )       
  ),
  
  column(8,
         renderPlot({
           
           d = input$dpop
           c = input$crit
           s <- seq(-3,3,0.01)
           
           n1 <- data.frame(x = s, y = dnorm(s, mean = 0, sd = 1))
           n2 <- data.frame(x = s, y = dnorm(s, mean = d, sd = 1))
           
           plot(n1$x, n1$y, type="l", xaxt = "n", yaxt = "n", ylab = "", xlab="", col="red4")
           lines(n2$x, n2$y, col="skyblue4")
           abline(v = c, lty=2)
           text(0, 0.01, "Noise", col="red4", cex = 1.25)
           text(d, 0.01, "Signal", col = "skyblue4", cex=1.25)
           text(c-.1, 0.2, "Criterion",srt=90, cex=1.25)
           
           
         })
  )
)
```


```{r echo = FALSE}
dprimesim <- function(dpop, c, trials, sims) {
  
  results <- vector()
  ntargets <- vector()
  d <- vector()
  
  #for all the n targets
  for (t in 1:(trials-1)) {
    
    #simulate 1000 people and get an average
    for (sub in 1:sims) {
      
      #sample values to represent responses
      nt <- trials - t
      nt_resp <- rnorm(nt,m=0,sd=1)
      t_resp <- rnorm(t,m=dpop,sd=1)
      
      #set criterion and get HIT and FA rates
      hr <- sum(t_resp >= c)/t
      far <- sum(nt_resp >= c)/nt
      
      #adjust if they are 0 or 1
      if (hr == 0) {
        hr = 1/(2*t)
      }
      
      if (hr == 1) {
        hr = 1 - 1/(2*t)
      }
      
      if (far == 0) {
        far = 1/(2*nt)
      }
      
      if (far == 1) {
        far = 1 - (1/(2*nt))
      }
      
      #calculate d'
      #store subject sim
      results[sub] = qnorm(hr) - qnorm(far)
    }
    
    #store results
    ntargets[t] = (t/trials)
    d[t] = mean(results)
  }
  
  return(data.frame("xvalues" = ntargets, "yvalues" = d))
}

```

```{r echo = FALSE}

#graph rate of targets (x-axis) and d' (y-axis)
renderPlot({
  
  sample <- dprimesim(input$dpop,input$crit,input$ntrials, as.numeric(input$nsim))
  
  plot(sample$xvalues, sample$yvalues, type = "b", pch = 19, 
       col = "skyblue4", xlab = "Target Rate (Number of Targets/Total Trials)", ylab = "Discriminability (d')", ylim=c(0,3))
  
  abline(h=input$dpop, lty=2)
})
```